---
title: "mapSpain CDN for CartoBase ANE"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 3
params:
  last_year: 2024
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  tidy = "styler",
  comment = "#>",
  fig.path = "img/README-",
  dev = "ragg_png",
  dpi = 300
)
```

![Data](https://img.shields.io/badge/last%20data%20available-%60r%20params$last_year%60-green)
![Last
Update](https://img.shields.io/badge/updated-%60r%20gsub(%22-%22,%22--%22,Sys.Date())%60-blue)

This branch is used as CDN to distribute CartoBase ANE data to **mapSpain**.

Due to poor API, data is not easily reachable, so this branch would be used as
API endpoint.

Last update : **`r Sys.Date()`**, containing **data up to
`r params$last_year`**.

## Source data

Atlas Nacional de España (ANE) [CC BY
4.0](https://creativecommons.org/licenses/by/4.0/deed.en)
[ign.es](http://www.ign.es).

<http://centrodedescargas.cnig.es/CentroDescargas/buscar.do?filtro.codFamilia=CAANE>

## Structure

Raw data is available under data-raw/YYYY. Zipped version and unzipped (selected
files) are provided.

## Distribution

The `dist` folder contains a distribution of [**CartoBase
ANE**](https://www.ign.es/web/ane-area-ane) on
[**GPKG**](https://en.wikipedia.org/wiki/GeoPackage) format.

```{r include=FALSE}
size <- file.size("dist/CartoBase.zip")
class(size) <- "object_size"
```

A `zip` file ([dist/CartoBase.zip](dist/CartoBase.zip),
`r format(size, units = "auto")` ) is also provided, including all the data on
`dist`.

## Files included

The files provided on `dist` come from the folder `todo` from each distribution.
These file contains the whole history on each dataset, with two relevant fields,
allowing to track historical changes between data sets:

-   `fecha_alta`: Start validity date
-   `fecha_baja`: End validity date

It is necessary to apply some filtering depending on the date requested:

```{r message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(ggplot2)

munic <- read_sf("dist/se89_3_admin_muni_a_x.gpkg", quiet = TRUE) %>%
  filter(rotulo == "Arenas del Rey")


ggplot(munic) +
  geom_sf() +
  facet_wrap(~fecha_alta, nrow = 1)


munic %>%
  st_drop_geometry() %>%
  select(fecha_alta, fecha_baja, rotulo) %>%
  knitr::kable()
```

An example of filtering:

```{r message=FALSE, warning=FALSE, out.width="33%", fig.show="hold", fig.asp=1.5}
# 2010 shape
year <- 2010
year.fmt <- as.Date(paste(year, "01", "01", sep = "-"))

shp2010 <- munic %>%
  # If fecha_baja is NA means it is the current valid shape
  mutate(fecha_baja = ifelse(is.na(fecha_baja), Sys.Date(), fecha_baja)) %>%
  filter(fecha_alta < year.fmt & year.fmt <= fecha_baja)

ggplot(shp2010) +
  geom_sf() +
  labs(title = year)

# 2017 shape
year <- 2017
year.fmt <- as.Date(paste(year, "01", "01", sep = "-"))

shp2017 <- munic %>%
  mutate(fecha_baja = ifelse(is.na(fecha_baja), Sys.Date(), fecha_baja)) %>%
  filter(fecha_alta < year.fmt & year.fmt <= fecha_baja)

ggplot(shp2017) +
  geom_sf() +
  labs(title = year)

# Current value

shpcurrent <- munic %>% filter(is.na(fecha_baja))


ggplot(shpcurrent) +
  geom_sf() +
  labs(title = "Current")
```

### Table of files

Full reference of each dataset in the
[data-raw](%60r%20paste0(%22data-raw/%22,params$last_year)%60) folder.

There are `r length(list.files("dist/", ".gpkg"))` files available:

```{r echo=FALSE}
library(dplyr)

files <- list.files("dist/", ".gpkg")


df <- data.frame(
  file = files,
  res = NA,
  stringsAsFactors = FALSE
)


if (length(grep("3", df$file)) > 0) {
  df[grep("3", df$file), ]$res <- "3"
}
if (length(grep("6m5", df$file) > 0)) {
  df[grep("6m5", df$file), ]$res <- "6.5"
}
if (length(grep("10", df$file)) > 0) {
  df[grep("10", df$file), ]$res <- "10"
}
if (length(grep("14", df$file)) > 0) {
  df[grep("14", df$file), ]$res <- "14"
}
if (length(grep("60", df$file)) > 0) {
  df[grep("60", df$file), ]$res <- "60"
}

df <- df %>% arrange(as.double(res), file)

df$scope <- ifelse(df$res <= 10, "Spain", "Europe")
df$scope <- ifelse(df$re == 60, "World", df$scope)

size <- c()

# Sizes
for (i in seq_len(nrow(df))) {
  s <- file.size(file.path("dist", df[i, 1]))
  class(s) <- "object_size"
  size <- c(size, format(s, units = "auto"))
}

df$size <- size


knitr::kable(df)
```

## Preview

```{r preview, echo=FALSE, fig.keep="all", message=FALSE, warning=FALSE, out.width="33%"}
par(mar = c(2, 2, 2, 2))
for (i in seq_len(nrow(df))) {
  path <- file.path("dist", df[i, "file"])

  input <- st_read(path, stringsAsFactors = FALSE, quiet = TRUE)


  plot(st_geometry(input), axes = TRUE, main = df[i, "file"])
}
```
