---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")

knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) {
      par(mar = c(0.1, 0.1, 1.3, 0.1))
    }
  }
)


```


# Introduction

[**`mapSpain`**](https://ropenspain.github.io/mapSpain/) is a package designed to provide geographical information of Spain at different levels.


**`mapSpain`** provides shapefiles of municipalities, provinces, autonomous communities
and NUTS levels of Spain. It also provides hexbin shapefiles and other complementary shapes, as the usual lines around the Canary Islands.

**`mapSpain`** provides access to map tiles of public organisms of Spain, that can be
represented on static maps via `mapSpain::esp_getTiles()` or on a **R** leaflet map using `mapSpain::addProviderEspTiles()`.

On top of that, **`mapSpain`** also has a powerful dictionary that translate provinces and other regions to English, Spanish, Catalan, Basque language or Galician, and also converts those names to different coding standards, as NUTS, ISO2 or the coding system used by the INE, that is the official statistic agency of Spain.


# Caching

`mapSpain` provides a dataset and tile caching capability, that could be set as:

```r
options(mapSpain_cache_dir = "~/path/to/dir")

OR 

options(gisco_cache_dir = "~/path/to/dir")
```

`mapSpain` relies on [`giscoR`](https://dieghernan.github.io/giscoR/) for downloading some files, and both packages are well synchronized, so if you already use `giscoR` and you have set your caching options for that package it would be recognized too by `mapSpain`.

# Installation

Install `mapSpain` from
[**CRAN**](https://CRAN.R-project.org/package=mapSpain):

``` r
install.packages(“mapSpain”)
```

For installing the development version on ([Github](https://github.com/rOpenSpain/mapSpain/)):

```r
library(remotes)
install_github("rOpenSpain/mapSpain")
```


# Basic example

Some examples of what `mapSpain`can do:

```{r basic, fig.width=7, fig.asp=1}
library(mapSpain)
library(sf)

country <- esp_get_country()
lines <- esp_get_can_box()

plot(st_geometry(country), axes = TRUE, col = "brown")
plot(lines, add = TRUE)


# Plot provinces

Andalucia <- esp_get_prov("Andalucia")
plot(
  st_geometry(Andalucia),
  col = "darkgreen",
  border = "white",
  axes = TRUE
)

# Plot municipalities

Euskadi_CCAA <- esp_get_ccaa("Euskadi")
Euskadi <- esp_get_munic(region = "Euskadi")

# Add layers
plot(
  st_geometry(Euskadi_CCAA),
  col = "grey50",
  axes = TRUE
)

plot(
  Euskadi[, "cpro"],
  pal = c("red2", "darkgreen", "ivory2"),
  add = TRUE,
  key.pos = NULL,
  border = "grey50",
  main = "Municipalities of the Basque Country"
)

```


# Thematic maps

This is an example on how `mapSpain` can be used to create thematic maps. For plotting purposes we would use the [`cartography` package](https://riatelab.github.io/cartography/docs/index.html), however any package that handles `sf` objects (e.g. `ggplot2`, `tmap`, `leaflet`, etc. could be used). 

```{r thematic}

# Population density of Spain - Center

pop <- mapSpain::pobmun19
Spain <- esp_get_country()
munic <- esp_get_munic()

# Get area (km2) - Use LAEA projection
municarea <- as.double(st_area(st_transform(munic, 3035)) / 1000000)
munic$area <- municarea

munic.pop <- merge(munic, pop, all.x = TRUE)

munic.pop$dens <- munic.pop$pob19 / munic.pop$area



library(cartography)

br <-
  c(0,
    10,
    25,
    100,
    200,
    500,
    1000,
    5000,
    10000,
    Inf)

pal <- hcl.colors(length(br) - 1, pal = "inferno", alpha = 0.7)

# Add layers


plot(st_geometry(Spain), col = pal[1], border = NA)

choroLayer(
  munic.pop,
  var = "dens",
  breaks = br,
  legend.pos = "n",
  col = pal,
  colNA = pal[1],
  border = NA,
  add = TRUE
)

plot(esp_get_can_box(), add = TRUE)


brlabs <- prettyNum(br, big.mark = ",")
brlabs[c(1, length(brlabs))] <- ""
legendChoro(
  pos = "topleft",
  breaks = brlabs,
  col = pal,
  title.txt = "Density Pop.\nkm2",
  nodata = FALSE
)




```

