---
title: "Working with imagery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with imagery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      warning = FALSE)

knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) {
      par(mar = c(0,0,0,0))
    }
  }
)

# Install reactable if not already installed
if (isFALSE(requireNamespace("reactable", quietly = TRUE))) {
  install.packages("reactable")
}

```

`mapSpain` provides a powerful interface for working with imagery. `mapSpain` can 
download static files as `.png` or`jpeg` files (depending on the Web Map Service) and
use them along your shapefiles.

`mapSpain` also includes a plugin for **R** [leaflet](https://rstudio.github.io/leaflet/) package, that allows you to include
several basemaps on your interactive maps.

The services are implemented via the leaflet plugin [leaflet-providersESP](dieghernan.github.io/leaflet-providersESP/). You can check a display of each provider on the previous link.

## Static tiles

An example of how you can include several tiles to create a static map. We focus here on the earthquakes of the last 365 days in the island of Tenerife. Note that the tile is actively maintained so the final plot based on a static file would vary depending on the date.

In this example, it is important to set `moveCAN = FALSE`, otherwise Tenerife would be
geographically displaced while the imagery won't.

```{r static1, fig.asp=1, warning=FALSE}
library(mapSpain)
library(sf)
library(cartography)

Tenerife <- esp_get_nuts(region = "Tenerife", moveCAN = FALSE)
Tenerife <- st_transform(Tenerife, 4083)


Base <- esp_getTiles(Tenerife, "MDT.Relieve", bbox_expand = 0.2)
Earthquake <-
  esp_getTiles(Tenerife, "Geofisica.Terremotos365dias", bbox_expand = 0.2)

tilesLayer(Base)
plot(st_geometry(Tenerife),
     col = adjustcolor("red", alpha.f = 0.4),
     add = TRUE)
tilesLayer(Earthquake, add = TRUE)

```

Static tiles could be loaded with or without an alpha value, that controls the transparency of the object:

```{r static2, fig.asp=1, warning=FALSE}

Madrid <- esp_get_ccaa("Madrid", epsg = 3857)

# Example without transparency

BG <- esp_getTiles(Madrid, "IGNBase.Gris")
ADIF.opaque <-
  esp_getTiles(Madrid, "RedTransporte.Ferroviario",
               transparent = FALSE)

tilesLayer(BG)
plot(
  st_geometry(Madrid),
  col = adjustcolor("red", alpha.f = 0.2),
  add = TRUE,
  border = NA
)
tilesLayer(ADIF.opaque, add = TRUE)

# Example with transparency

ADIF.alpha <- esp_getTiles(Madrid,
                           "RedTransporte.Ferroviario")

tilesLayer(BG)
plot(
  st_geometry(Madrid),
  col = adjustcolor("red", alpha.f = 0.2),
  add = TRUE,
  border = NA
)
tilesLayer(ADIF.alpha, add = TRUE)

```

Another nice feature is the ability of masking the tiles, so more advanced maps can be plotted:

```{r static3, fig.width= 7.3, fig.height=3.7, warning=FALSE}

Asturias <- esp_get_ccaa("Asturias", epsg = 3857)

Basemap <- esp_getTiles(Asturias, "IDErioja", bbox_expand = 0.3)

Mask <- esp_getTiles(Asturias, "LiDAR", mask = TRUE)


tilesLayer(Basemap)
tilesLayer(Mask, add = TRUE)

```

## Dynamic maps with Leaflet


`mapSpain` provides a plugin to be used with the `leaflet` package. Here you can find 
the leaflet versions of some of the previous examples:


```{r leaflet1}

library(leaflet)

Tenerife.leaf <-
  esp_get_nuts(region = "Tenerife",
               epsg = 4326,
               moveCAN = FALSE)


bbox <-
  as.double(round(st_bbox(Tenerife.leaf) + c(-1, -1, 1, 1), 2))

# Start leaflet
m <- leaflet(Tenerife.leaf,
             width = "100%",
             options = leafletOptions(minZoom = 9,
                                      maxZoom = 18))

# Add layers
m <- m %>%
  addProviderEspTiles("MDT.Relieve") %>%
  addPolygons(color = NA,
              fillColor = "red",
              group = "Polygon") %>%
  addProviderEspTiles("Geofisica.Terremotos365dias",
                      group = "Earthquakes")

# Add additional options
m  %>%
  addLayersControl(
    overlayGroups = c("Polygon", "Earthquakes"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% setMaxBounds(bbox[1], bbox[2], bbox[3], bbox[4])


```


```{r leaflet2}

Madrid.leaflet <- esp_get_ccaa("Madrid", epsg = 4326)

m2 <- leaflet(Madrid.leaflet, width = "100%") %>%
  addProviderEspTiles("IGNBase.Gris") %>%
  addPolygons(color = NA,
              fillColor = "red",
              group = "Polygon") %>%
  addProviderEspTiles("RedTransporte.Ferroviario",
                      group = "Railway")

# Add layer control

m2 %>% addLayersControl(
  overlayGroups = c("Polygon", "Railway"),
  options = layersControlOptions(collapsed = FALSE)
)

```

## Available providers 

The data frame `leaflet.providersESP.df` includes the data of the available providers 
you can use on functions described above. This data frame includes all the
parameters needed to replicate the API request. In this summary only the
parameter `url_static` is shown:

```{r providers, echo=FALSE}

library(reactable)

df <- leaflet.providersESP.df[leaflet.providersESP.df$field == "url_static", ]

reactable(df,
          filterable = TRUE,
          searchable = TRUE,
          showPageSizeOptions = TRUE,
          striped = TRUE,
          paginationType = "jump"
          )

```

## Session Info

<details>
  <summary><strong>Details</strong></summary>
```{r session_info, echo=FALSE}
sessionInfo()

```
</details>

