---
title: "Working with imagery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with imagery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  out.width = "100%"
)

# Install reactable if not already installed
if (isFALSE(requireNamespace("reactable", quietly = TRUE))) {
  install.packages("reactable")
}

```

`mapSpain` provides a powerful interface for working with imagery. `mapSpain` 
can download static files as `.png` or`jpeg` files (depending on the Web Map 
Service) and use them along your shapefiles.

`mapSpain` also includes a plugin for **R** 
[leaflet](https://rstudio.github.io/leaflet/) package, that allows you to 
include several basemaps on your interactive maps.

The services are implemented via the leaflet plugin [leaflet-providersESP](dieghernan.github.io/leaflet-providersESP/). You can 
check a display of each provider on the previous link.

## Static tiles

An example of how you can include several tiles to create a static map. We 
focus here on layer provided by La Rioja's [Infraestructura de Datos Espaciales (IDERioja)]https://www.iderioja.larioja.org/

**When working with imagery, it is important to set `moveCAN = FALSE`, 
otherwise the images for the Canary Islands won't be accurate.**

```{r static1, fig.asp=1}

library(mapSpain)
library(sf)
library(tmap)

# Logroño

LGN <- esp_get_munic(munic = "Logroño")

# Convert to Mercator (EPSG:3857) as a general advice when working with tiles
LGN <- st_transform(LGN, 3857)

tile_LGN <- esp_getTiles(LGN, "IDErioja", bbox_expand = 0.5)

tm_shape(tile_LGN, raster.downsample = FALSE) +
  tm_rgb() +
  tm_shape(LGN) +
  tm_borders(lwd = 2, lty = "dotted")


```

Some tiles could be loaded with or without an alpha value, that controls the 
transparency of the object:

```{r static2, fig.asp=1}

Madrid <- esp_get_ccaa("Madrid", epsg = 3857)

# Example without transparency

BG <- esp_getTiles(Madrid, "IGNBase.Gris")
ADIF.opaque <-
  esp_getTiles(Madrid, "RedTransporte.Ferroviario",
               transparent = FALSE)

tm_shape(BG, raster.downsample = FALSE) +
  tm_rgb() +
  tm_shape(ADIF.opaque, raster.downsample = FALSE) +
  tm_rgb()

# Example with transparency

ADIF.alpha <- esp_getTiles(Madrid,
                           "RedTransporte.Ferroviario",
                           transparent = TRUE)

# Same code than above for plotting

tm_shape(BG) +
  tm_rgb() +
  tm_shape(ADIF.alpha) +
  tm_rgb()

```

Another nice feature is the ability of masking the tiles, so more advanced maps can be plotted:

```{r static3}

Marbella <- esp_get_munic(munic = "Marbella")
Marbella <- st_transform(Marbella, 3857)

Basemap <- esp_getTiles(Marbella, "PNOA", bbox_expand = 0.1, )

Mask <- esp_getTiles(Marbella, "IGNBase.Gris", mask = TRUE)


tm_shape(Basemap, raster.downsample = FALSE) +
  tm_rgb() +
  tm_shape(Mask, raster.downsample = FALSE) +
  tm_rgb()

```

## Dynamic maps with Leaflet


`mapSpain` provides a plugin to be used with the `leaflet` package. Here you 
can find some quick examples:


```{r leaflet1}

library(leaflet)

Tenerife.leaf <-
  esp_get_nuts(region = "Tenerife",
               epsg = 4326,
               moveCAN = FALSE)


bbox <-
  as.double(round(st_bbox(Tenerife.leaf) + c(-1, -1, 1, 1), 2))

# Start leaflet
m <- leaflet(Tenerife.leaf,
             width = "100%",
             options = leafletOptions(minZoom = 9,
                                      maxZoom = 18))

# Add layers
m <- m %>%
  addProviderEspTiles("MDT.Relieve") %>%
  addPolygons(color = NA,
              fillColor = "red",
              group = "Polygon") %>%
  addProviderEspTiles("Geofisica.Terremotos365dias",
                      group = "Earthquakes")

# Add additional options
m  %>%
  addLayersControl(
    overlayGroups = c("Polygon", "Earthquakes"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% setMaxBounds(bbox[1], bbox[2], bbox[3], bbox[4])


```


```{r leaflet2}

Madrid.leaflet <- esp_get_ccaa("Madrid", epsg = 4326)

m2 <- leaflet(Madrid.leaflet, width = "100%") %>%
  addProviderEspTiles("IGNBase.Gris") %>%
  addPolygons(color = NA,
              fillColor = "red",
              group = "Polygon") %>%
  addProviderEspTiles("RedTransporte.Ferroviario",
                      group = "Railway")

# Add layer control

m2 %>% addLayersControl(
  overlayGroups = c("Polygon", "Railway"),
  options = layersControlOptions(collapsed = FALSE)
)

```

## Available providers 

The data frame `leaflet.providersESP.df` includes the data of the available providers 
you can use on functions described above. This data frame includes all the
parameters needed to replicate the API request. In this summary only the
parameter `url_static` is shown:

```{r providers, echo=FALSE}

library(reactable)

df <- leaflet.providersESP.df[leaflet.providersESP.df$field == "url_static", ]

reactable(df,
          filterable = TRUE,
          searchable = TRUE,
          showPageSizeOptions = TRUE,
          striped = TRUE,
          paginationType = "jump"
          )

```

## Session Info

<details>
  <summary><strong>Details</strong></summary>
```{r session_info, echo=FALSE}
sessionInfo()

```
</details>

