---
title: "mapSpain: maps of Spain in R"
author:
  - name: Diego Hernangómez
    orcid: 0000-0001-8457-4658
description: Create maps of Spain in R quickly and simply.
toc: true
toc-depth: 2
lang: en
language:
  crossref-fig-title: "Figure"
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    warning: false
    message: false
    dpi: 300
    tidy: "styler"
    dev: "ragg_png"
    out.width: "100%"
---

## Welcome to mapSpain

### Motivation

**mapSpain** helps you create maps for the main administrative levels of Spain.
It also supports imagery from WMS and WMTS services, either as georeferenced
static rasters or as dynamic layers in Leaflet maps.

The package also includes helpers to normalize Autonomous Community and
province names. These helpers make it easier to join, clean and transform data,
whether or not the data is spatial.

The main data sources used by **mapSpain** are:

- [GISCO](https://ec.europa.eu/eurostat/web/gisco) (Eurostat), through the
  [**giscoR**](https://ropengov.github.io/giscoR/) package.
- [Instituto Geográfico Nacional](https://www.ign.es/) (IGN).
- Spanish public institutions that publish WMTS and WMS tile services
  (<https://www.idee.es/web/idee/segun-tipo-de-servicio>).

Functions return `sf` objects from the **sf** package or `SpatRaster` objects
from the **terra** package.

Package website: <https://ropenspain.github.io/mapSpain/>.

### Installation

#### CRAN

```{r}
#| label: instcran
#| eval: false
install.packages("mapSpain", dependencies = TRUE)
```

#### Development version

Use [r-universe](https://ropenspain.r-universe.dev/ui#builds):

```{r}
#| label: instruniverse
#| eval: false
# Enable this universe.
install.packages(
  "mapSpain",
  repos = c(
    "https://ropenspain.r-universe.dev",
    "https://cloud.r-project.org"
  ),
  dependencies = TRUE
)
```

#### Installation with pak from GitHub

```{r}
#| label: instremotes
#| eval: false
install.packages("pak")
pak::pak("rOpenSpain/mapSpain", dependencies = TRUE)
```

### A quick example

```{r}
#| label: fig-intro
#| fig-cap: Municipalities of Galicia
library(mapSpain)
library(tidyverse)

galicia <- esp_get_munic_siane(region = "Galicia") |>
  # Standardize labels.
  mutate(Provincia = esp_dict_translate(ine.prov.name, "es"))

ggplot(galicia) +
  geom_sf(aes(fill = Provincia), color = "grey70") +
  labs(title = "Provinces of Galicia") +
  scale_fill_discrete(type = hcl.colors(4, "Blues")) +
  theme_bw()
```

You can also inspect the dataset interactively:

```{r}
#| label: galicia_data
library(reactable)

reactable(
  galicia,
  searchable = TRUE,
  striped = TRUE,
  filterable = TRUE,
  height = 350,
  elementId = "munis-galicia"
)
```

### Comparing mapSpain with other alternatives

The next example compares **mapSpain** with other packages that provide `sf` or
`SpatVector` objects for country boundaries.

```{r}
#| label: fig-compara
#| fig-cap: |
#|   Comparison of boundary resolution across map sources. The boundaries are
#|   shown over an orthophoto of the Ferrol estuary.
library(sf) # Spatial data manipulation.

# rnaturalearth
library(rnaturalearth)
esp_rnat <- ne_countries("large", country = "Spain", returnclass = "sf") |>
  st_transform(3857)

# mapSpain
esp_mapspain <- esp_get_spain(epsg = 4326) |>
  st_transform(3857)

# geodata (GADM)
library(geodata)
esp_geodata <- geodata::gadm("ES", path = ".", level = 0) |>
  # Convert from SpatVector to an sf object.
  sf::st_as_sf() |>
  st_transform(3857)

# geobounds
library(geobounds)
esp_geobounds <- geobounds::gb_get_adm0("ESP", cache_dir = ".") |>
  st_transform(3857)

# Image of the Ferrol estuary.
tile <- esp_get_munic_siane(munic = "Ferrol", epsg = 3857) |>
  esp_get_tiles("PNOA", bbox_expand = 0.5, zoommin = 1)

# Prepare the plot.
library(tidyterra)

esp_all <- bind_rows(esp_rnat, esp_mapspain, esp_geodata, esp_geobounds)

esp_all$source <- as_factor(c(
  "rnaturalearth",
  "mapSpain",
  "geodata (GADM)",
  "geoBoundaries"
))

ggplot(esp_all) +
  geom_spatraster_rgb(data = tile, maxcell = Inf, alpha = 0.5) +
  geom_sf(
    aes(color = source, linetype = source),
    fill = NA,
    show.legend = "line",
    linewidth = 0.5,
    alpha = 0.7
  ) +
  coord_sf(
    crs = 4326,
    xlim = c(-8.384421, -8.154413),
    ylim = c(43.43201, 43.59545),
    expand = FALSE
  ) +
  scale_color_manual(values = c("red", "black", "blue", "orange")) +
  scale_linetype_manual(values = c("solid", "dotted", "dashed", "twodash")) +
  theme_void() +
  labs(title = "Ferrol estuary")
```

- **rnaturalearth**: less precise boundary interpretation.
- **mapSpain**: satisfactory results.
- **GADM** (through **geodata**): very precise results.
- **geoBoundaries** (through **geobounds**): satisfactory results.

### Caching

**mapSpain** uses web resources. By default, downloaded files are stored in
`tempdir()` for reuse during the current session.

Use `esp_set_cache_dir()` to choose a user-specific download directory. Add
`install = TRUE` to make that cache configuration persistent across sessions.

``` r
esp_set_cache_dir("~/R/mapslib/mapSpain", install = TRUE, verbose = TRUE)

#> ℹ mapSpain cache directory is C:/Users/XXXX/Documents/R/mapslib/mapSpain.

munic <- esp_get_munic_siane(verbose = TRUE)

#> ℹ Cache directory is C:/Users/XXXX/Documents/R/mapslib/mapSpain/siane.
#> ✔ File already cached: C:/Users/XXXX/Documents/R/mapslib/mapSpain/siane/se89_3_admin_muni_a_x.gpkg.
#> ℹ Cache directory is C:/Users/diego/Documents/R/mapslib/GISCO/siane.
#> ✔ File already cached: C:/Users/XXXX/Documents/R/mapslib/mapSpain/siane/se89_3_admin_muni_a_y.gpkg
```

## Dictionary

### Functions for working with text

**mapSpain** provides two related functions for working with names and codes:

- `esp_dict_region_code()` converts text labels into Autonomous Community and
  province codes. Supported coding schemes are ISO2, NUTS and INE codes
  (`codauto` and `cpro`).
- `esp_dict_translate()` translates text into Spanish, English, Catalan,
  Galician or Basque.

These functions are also useful outside spatial workflows, for example when you
need to standardize Autonomous Community and province codes in ISCIII COVID
data.

#### `esp_dict_region_code()`

```{r}
vals <- c("Errioxa", "Coruna", "Gerona", "Madrid")

esp_dict_region_code(vals, destination = "nuts")
esp_dict_region_code(vals, destination = "cpro")
esp_dict_region_code(vals, destination = "iso2")

# From ISO to other codes.

iso2vals <- c("ES-M", "ES-S", "ES-SG")
esp_dict_region_code(iso2vals, origin = "iso2")

iso2vals <- c("ES-GA", "ES-CT", "ES-PV")

esp_dict_region_code(iso2vals, origin = "iso2", destination = "nuts")

# Support mixed levels.
valsmix <- c("Centro", "Andalucia", "Seville", "Menorca")
esp_dict_region_code(valsmix, destination = "nuts")

esp_dict_region_code(c("Murcia", "Las Palmas", "Aragón"), destination = "iso2")
```

#### `esp_dict_translate()`

```{r}
vals <- c("La Rioja", "Sevilla", "Madrid", "Jaen", "Orense", "Baleares")

esp_dict_translate(vals, lang = "en")
esp_dict_translate(vals, lang = "es")
esp_dict_translate(vals, lang = "ca")


esp_dict_translate(vals, lang = "eu")
esp_dict_translate(vals, lang = "ga")
```

## Political boundaries

**mapSpain** includes functions to retrieve political boundaries at several
levels:

- Whole country.
- [NUTS](https://ec.europa.eu/eurostat/web/nuts/background) (Eurostat).
  Eurostat statistical classification, with levels 0 (country), 1, 2
  (Autonomous Communities) and 3.
- Autonomous Communities.
- Provinces.
- Municipalities.

For Autonomous Communities, provinces and municipalities, there are two
families of functions: `esp_get_xxxx()` for GISCO data and
`esp_get_xxxx_siane()` for IGN data.

The information is available in different projections and resolution levels.

```{r}
#| label: fig-pais
#| fig-cap: Map of Spain
esp <- esp_get_spain_siane(moveCAN = FALSE)

ggplot(esp) +
  geom_sf(fill = "#f9cd94") +
  theme_light()
```

### The Canary Islands case

By default, most **mapSpain** functions move the Canary Islands closer to the
mainland to improve visualization. Disable this behavior with `moveCAN = FALSE`
when you need geometries in their original position.

The package also provides helpers for drawing boxes around the inset map. See
the
[examples](https://ropenspain.github.io/mapSpain/reference/esp_get_can_box.html#examples)
in the reference page.

```{r}
#| label: fig-can
#| fig-cap: Map of Spain with displaced Canary Islands
esp_can <- esp_get_spain()
can_prov <- esp_get_can_provinces()
can_box <- esp_get_can_box()

ggplot(esp_can) +
  geom_sf() +
  geom_sf(data = can_prov) +
  geom_sf(data = can_box)
```

**Use `moveCAN = FALSE` when working with imagery, interactive maps or spatial
analysis.**

### NUTS

```{r}
#| label: fig-nuts
#| fig-cap: NUTS 1 regions of Spain
nuts1 <- esp_get_nuts(resolution = 20, epsg = 3035, nuts_level = 1)

ggplot(nuts1) +
  geom_sf() +
  labs(title = "NUTS1: low resolution")
```

```{r}
#| label: fig-nuts3
#| fig-cap: NUTS 3 regions of Spain
# Balearic Islands NUTS3.
nuts3_baleares <- c("ES531", "ES532", "ES533")
paste(esp_dict_region_code(nuts3_baleares, "nuts"), collapse = ", ")

nuts3_sf <- esp_get_nuts(region = nuts3_baleares)

ggplot(nuts3_sf) +
  geom_sf(aes(fill = NAME_LATN)) +
  labs(fill = "Balearic Islands: NUTS3") +
  scale_fill_viridis_d() +
  theme_minimal()
```

### Autonomous Communities

```{r}
#| label: fig-ccaa
#| fig-cap: "Autonomous Communities of Spain"
ccaa <- esp_get_ccaa(
  ccaa = c(
    "Catalunya",
    "Comunidad Valenciana",
    "Aragón",
    "Baleares"
  )
)

ccaa <- ccaa |>
  mutate(ccaa_cat = esp_dict_translate(ine.ccaa.name, "ca"))

ggplot(ccaa) +
  geom_sf(aes(fill = ccaa_cat)) +
  labs(fill = "Autonomous Communities") +
  theme_minimal() +
  scale_fill_discrete(type = hcl.colors(4, "Plasma"))
```

### Provinces using the `*_siane` version

Passing a higher-level entity, such as Andalusia, returns all provinces within
that entity.

```{r}
#| label: fig-prov
#| fig-cap: "Extracting multiple provinces through Autonomous Communities"
provs <- esp_get_prov_siane(c(
  "Andalucía",
  "Ciudad Real",
  "Murcia",
  "Ceuta",
  "Melilla"
))

ggplot(provs) +
  geom_sf(aes(fill = prov.shortname.es), alpha = 0.9) +
  scale_fill_discrete(type = hcl.colors(12, "Cividis")) +
  theme_minimal() +
  labs(fill = "Provinces")
```

### Municipalities

```{r}
#| label: fig-munic
#| fig-cap: "Extracting municipalities"
munic <- esp_get_munic(region = "Segovia") |>
  # Example data: INE population.
  left_join(
    mapSpain::pobmun25 |>
      select(-name),
    by = c("cpro", "cmun")
  )

ggplot(munic) +
  geom_sf(aes(fill = pob25), alpha = 0.9, color = NA) +
  scale_fill_gradientn(
    colors = hcl.colors(100, "Inferno"),
    n.breaks = 10,
    labels = scales::label_comma(),
    guide = guide_legend()
  ) +
  labs(
    fill = "Population",
    title = "Population in Segovia",
    subtitle = "INE data (2025)"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect("grey80"),
    text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### Hexbin maps

Hexbin maps are available as squares and hexagons for provinces and Autonomous
Communities.

```{r}
#| label: fig-hex
#| layout-ncol: 2
#| fig-cap: Hexbin maps with mapSpain
#| fig-subcap:
#|   - Squares
#|   - Hexagons
cuad <- esp_get_hex_ccaa()
hex <- esp_get_grid_prov()

ggplot(cuad) +
  geom_sf() +
  geom_sf_text(aes(label = iso2.ccaa.code)) +
  theme_void()


ggplot(hex) +
  geom_sf() +
  geom_sf_text(aes(label = iso2.prov.code)) +
  theme_void()
```

## Imagery

**mapSpain** can also use map imagery, such as satellite imagery, basemaps and
roads, provided by different public institutions
(<https://www.idee.es/web/idee/segun-tipo-de-servicio>).

These images can be used to create static maps, as 3- or 4-band raster layers,
or as backgrounds for dynamic maps through the **leaflet** package.

The providers are taken from the **leaflet**
[leaflet-providersESP](https://dieghernan.github.io/leaflet-providersESP/)
plugin.

### Creating static maps

Several options are available for composing basemaps:

```{r}
#| label: fig-imagesestaticos
#| fig-cap:
#|   - Basemap extraction
#|   - Basemap with a mask
madrid_munis <- esp_get_munic_siane(region = "Madrid", epsg = 3857)
base_pnoa <- esp_get_tiles(madrid_munis, "PNOA", bbox_expand = 0.1, zoommin = 1)

library(tidyterra)

ggplot() +
  geom_spatraster_rgb(data = base_pnoa) +
  geom_sf(
    data = madrid_munis,
    color = "blue",
    fill = "blue",
    alpha = 0.25,
    linewidth = 0.5
  ) +
  theme_minimal() +
  labs(title = "Municipalities in Madrid")


# Use the `mask` option.
madrid <- esp_get_munic_siane(munic = "^Madrid$", epsg = 3857)

madrid_mask <- esp_get_tiles(
  madrid,
  "IDErioja.Claro",
  mask = TRUE,
  crop = TRUE,
  zoommin = 2
)

ggplot() +
  geom_spatraster_rgb(data = madrid_mask) +
  theme_void() +
  labs(
    title = "Basemap of Madrid",
    caption = "CC BY 4.0 www.iderioja.org"
  )
```

### Dynamic maps with mapSpain

Imagery layers can be used as backgrounds in static and interactive maps.

```{r}
#| label: leaflet
stations <- esp_get_railway(spatialtype = "point", epsg = 4326)

library(leaflet)
# Create an icon.

iconurl <- "https://ropenspain.github.io/mapSpain/icons/train.png"

train_icon <- makeIcon(iconurl, iconurl, 18, 18)

leaflet(stations, elementId = "railway", width = "100%", height = "60vh") |>
  addProviderEspTiles("IDErioja.Claro", group = "Base") |>
  addProviderEspTiles("MTN", group = "MTN") |>
  addProviderEspTiles("RedTransporte.Carreteras", group = "Roads") |>
  addProviderEspTiles(
    "RedTransporte.Ferroviario",
    group = "Railway lines"
  ) |>
  addMarkers(
    icon = train_icon,
    group = "Stations",
    popup = sprintf(
      "<strong>%s</strong>",
      stations$rotulo
    ) |>
      lapply(htmltools::HTML)
  ) |>
  addLayersControl(
    baseGroups = c("Base", "MTN"),
    overlayGroups = c("Stations", "Railway lines", "Roads"),
    options = layersControlOptions(collapsed = FALSE)
  ) |>
  hideGroup(c("Railway lines", "Roads"))
```

## Other resources

**mapSpain** includes additional
[functions](https://ropenspain.github.io/mapSpain/reference/index.html#section-natural)
for retrieving elevation, rivers and river basin data for Spain, as well as
Spanish
[transport infrastructure](https://ropenspain.github.io/mapSpain/reference/index.html#section-transport-infrastructure)
lines and points, such as roads and railway lines.
